---
title: "Lab1"
author: "Andrew Mitchell"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Lab1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(Lab1Intro)
```

# Links to Relevant Material

<GitHub Repository>[Andrew Mitchell GitHub Repository](https://github.com/ajmitchell715/Lab1Intro)

<R Packages Tutorial>[R Packages Tutorial](http://r-pkgs.had.co.nz/intro.html)

# Formulae

The first three of these formulae will be used to create:

* A mean vector for the variables in the data set

* A covariance matrix for the variables

* A correlation coefficient matrix for the variables

The other formulae listed describe relationships between the matrices but are not used for the purposes of creating the vector and matrices.

##  Formula 1-1: The mean of a variable k

$\bar{x_k}=\dfrac{1}{n}\Sigma_{j=1}^{n}x_{jk}$ for $k=1,2,...,p$

## Formula 1-4: The covariance of two variables i and k

$s_{ik}=\dfrac{1}{n}\Sigma_{j=1}^n(x_{ji}-\bar{x_i})(x_{jk}-\bar{x_k})$

## Formula 1-5: The correlation coefficient of two variables i and k

$$
\begin{eqnarray}
r_{ik} &=& \dfrac{s_{ik}}{\sqrt{s_{ii}}\sqrt{s_{kk}}}\\
&=& \dfrac{\Sigma_{j=1}^n(x_{ji}-\bar{x_i})(x_{jk}-\bar{x_k})}{\sqrt{\Sigma_{j=1}^n(x_{ji}-\bar{x_i})^2}\sqrt{\Sigma_{j=1}^n(x_{jk}-\bar{x_k})^2}}
\end{eqnarray}
$$

## Formulas 3-27: The mean and covariance matrices in terms of the data set

$\bar{\textbf{x}}=\dfrac{1}{n}\textbf{X}'1$

$\textbf{S}=\dfrac{1}{n-1}\textbf{X}'(\textbf{I}-\dfrac{1}{n}\textbf{11}')\textbf{X}$

## Formula 3-28: The sample standard deviation matrix

$$
\textbf{D}^{\dfrac{1}{2}}=\begin{bmatrix}
\sqrt{s_{11}} & 0 & \cdots & 0\\
0 & \sqrt{s_{22}} & \cdots & 0\\
\vdots & \vdots & \ddots & \vdots\\
0 & 0 & \cdots & \sqrt{s_{pp}}
\end{bmatrix}
$$

## Formula 3-29: The correlation coefficient matrix in terms of the sample standard deviation matrix and covariance matrix

$\textbf{R}=\textbf{D}^{\dfrac{-1}{2}}\textbf{SD}^{\dfrac{-1}{2}}$

# Functions Calculating the Mean Vector and Covariance and Correlation Matrices

## Creating the Data Set

The data used for this laboratory describe 41 machines' density and strength, with the latter two variables in the data set combining to describe strength.

```{r}
machine=read.table("T1-2.DAT")
colnames(machine)=c("Density","Machine Direction","Cross Direction")
```

## Creating the Mean Vector

```{r}
meanVec(machine)
```

## Creating the Covariance Matrix

```{r}
covMat(machine)
```

## Creating the Correlation Coefficient Matrix

```{r}
corMat(machine)
```

# Assessment for Students in MATH 4793

Clicker Quizzes: 10%

4 Assignments: 20%

Laboratories: 10%

2 Exams: 10% each

2 Projects: 10%

Final Exam: 30%
